---
title: "Using mixed models to study leaf-cutting ant foraging"
author: "Sasha Mikheyev and Manasee Weerathunga"
output: html_document
---


```{r message=FALSE}
library(tidyverse)
library(ape)
library(tidybayes)
library(brms)
library(ggtree)
library(sjPlot)
```

## Mixed model

```{r full_model, cache = T}
speciesdat <- read_csv("completedataR.csv") 
tre <- read.nexus("treenew.nexus") #Output from Phylomatic

inv.phylo <- MCMCglmm::inverseA(tre, nodes = "TIPS", scale = FALSE)
A <- solve(inv.phylo$Ainv)
rownames(A) <- rownames(inv.phylo$Ainv)
priorscomplete <- get_prior(Foraged ~ DBH + Distance + (1 | Colony) + (1 | Species) + (1 | TAG), data = speciesdat, family = bernoulli())
complete <- brm(Foraged ~ DBH + Distance + (1 | Colony) + (1 | Species) + (1 | TAG), data = speciesdat, family = bernoulli, cov_ranef = list(phylo = A), save_all_pars = TRUE, cores = 4, control = list(adapt_delta = 0.9), prior = priorscomplete, iter = 4000)
summary(complete)
plot(complete)
<<<<<<< .merge_file_kGwCd3
r2 <- list()
(r2[["complete"]] <- bayes_R2(complete))
=======
bayes_R2(complete)
tab_model(complete, show.intercept = FALSE, show.r2 = TRUE)
>>>>>>> .merge_file_uzGwX0
```

## Reduced models

Compute R2 for reduced models to estimate how big individual effects are

```{r cache=T, message=F}
# colony
form <- formula(Foraged ~  (1 | Colony))
priors <- get_prior(form, data = speciesdat, family = bernoulli())
reduced <- brm(form, data = speciesdat, family = bernoulli, cov_ranef = list(phylo = A), cores = 4, control = list(adapt_delta = 0.9), prior = priors, iter = 4000)
r2[["colony"]] <- bayes_R2(reduced)

# species effect
form <- formula(Foraged ~ (1 | Species) )
priors <- get_prior(form, data = speciesdat, family = bernoulli())
reduced <- brm(form, data = speciesdat, family = bernoulli, cov_ranef = list(phylo = A), cores = 4, control = list(adapt_delta = 0.9), prior = priors, iter = 4000)
r2[["species"]] <- bayes_R2(reduced)

# tag effect
form <- formula(Foraged ~  (1 | TAG) )
priors <- get_prior(form, data = speciesdat, family = bernoulli())
reduced <- brm(form, data = speciesdat, family = bernoulli, cov_ranef = list(phylo = A), cores = 4, control = list(adapt_delta = 0.9), prior = priors, iter = 4000)
r2[["tag"]] <- bayes_R2(reduced)

# DBH effect
form <- formula(Foraged ~  DBH)
priors <- get_prior(form, data = speciesdat, family = bernoulli())
reduced <- brm(form, data = speciesdat, family = bernoulli, cov_ranef = list(phylo = A), cores = 4, control = list(adapt_delta = 0.9), prior = priors, iter = 4000)
r2[["dbh"]] <- bayes_R2(reduced)

# distance effect
form <- formula(Foraged ~ Distance)
priors <- get_prior(form, data = speciesdat, family = bernoulli())
reduced <- brm(form, data = speciesdat, family = bernoulli, cov_ranef = list(phylo = A), cores = 4, control = list(adapt_delta = 0.9), prior = priors, iter = 4000)
r2[["distance"]] <- bayes_R2(reduced)

r2
```


## phylogenetic effect

We are are also testing the hypothesis that there is no phylogenetic effect (ignoring all other factors), following [Nakagawa and Schielzeth's (2012)](https://besjournals.onlinelibrary.wiley.com/doi/full/10.1111/j.2041-210x.2012.00261.x%4010.1111/%28ISSN%292041-210X.STATSTOO) Table 2 for distribution-specific variance.

```{r reduced_models, cache = TRUE}
phylo_only <- brm(Foraged ~ (1 | Species), data = speciesdat, family = bernoulli, cov_ranef = list(phylo = A), cores = 4, control = list(adapt_delta = 0.9))

bayes_R2(phylo_only)

hyp <- paste("sd_Species__Intercept /", "(sd_Species__Intercept + 3.141593^(2/3)) = 0")
(hyp <- hypothesis(phylo_only, hyp, class = NULL))
plot(hyp)
```

We find that the complete model has the best fit to the data

## Palatability coefficients

UsingWe can look a the coefficients of individual species, which correspond to how likely an they are to be foraged, aftera accounting for all the otehr factors, like the phylogenetic effect.

```{r plot_palatability}
palatcoef <- complete %>% spread_draws(b_Intercept, r_Species[Species ]) %>% group_by(Species) %>% 
  median_qi(Species_mean =  r_Species) %>% mutate(Species = gsub("_"," ", Species))
ggplot(palatcoef, aes(Species_mean)) +geom_histogram() + theme_minimal() + xlab("Palatability coefficients") + ylab("Number of tree species")
#ggsave("figures/palatbility.pdf", height = 5, width = 5)
```

```{r}
plotdat <- as.data.frame(speciesdat %>% group_by(Species) %>% summarize(Foraged = as.factor(max(Foraged))) %>% ungroup() %>% mutate(Species = gsub("_", " ", Species)) %>% select(Species, Foraged) %>% na.omit() %>% left_join(palatcoef %>% select(Species, Species_mean)))
row.names(plotdat) <- plotdat$Species

t2 <- tre 
t2$tip.label <- gsub("_", " ", t2$tip.label)

p1 <- ggtree(t2, layout = 'fan') %<+% plotdat + geom_tiplab2(aes(label=paste0('italic(\'', label, '\')'), color=Foraged), size = 2, offset=0.5, parse = TRUE) + scale_color_manual(values = c("black", "red"))
p2 <- gheatmap(p1, plotdat[,"Species_mean", drop=FALSE], offset = -1, width=0.05, colnames=FALSE, colnames_position = "bottom") + scale_fill_gradient2(low = "black", mid = "white", high = "red", midpoint = 0) + theme(legend.position = "bottom") 
print(p2)
#ggsave("figures/tree.pdf", height = 8, width = 8)
```
