---
title: "Using mixed models to study leaf-cutting ant foraging"
author: "Sasha Mikheyev and Manasee Weerathunga"
output: html_document
---


```{r}
library(tidyverse)
library(ape)
library(tidybayes)
library(brms)
library(ggtree)
```

# Mixed model

```{r full_model, cache = T}
speciesdat <- read_csv("completedataR.csv") %>% transmute(Species = gsub(" ", "_", Species), DBH =  DBH, Colony = Colony, Foraged = Foraged, Distance = Distance, Succession = Succession) %>% dplyr::distinct()
tre <- read.nexus("treenew.nexus") #Output from Phylomatic

inv.phylo <- MCMCglmm::inverseA(tre, nodes = "TIPS", scale = FALSE)
A <- solve(inv.phylo$Ainv)
rownames(A) <- rownames(inv.phylo$Ainv)
priorscomplete <- get_prior(Foraged ~ DBH + Distance + (1 | Colony) + (1 | Species), data = speciesdat, family = bernoulli())
priorscomplete$prior[1:3] <- "normal(0,10)" # make priors even flatter
priorscomplete$prior[6:9] <- "normal(0,10)"
complete <- brm(Foraged ~ DBH + Distance + (1 | Colony) + (1 | Species), data = speciesdat, family = bernoulli, cov_ranef = list(phylo = A), save_all_pars = TRUE, cores = 4, control = list(adapt_delta = 0.9), prior = priorscomplete)
summary(complete)
plot(complete)
bayes_R2(complete)
```

## phylogenetic effect

We are are also testing the hypothesis that there is no phylogenetic effect (ignoring all other factors), following [Nakagawa and Schielzeth's (2012)](https://besjournals.onlinelibrary.wiley.com/doi/full/10.1111/j.2041-210x.2012.00261.x%4010.1111/%28ISSN%292041-210X.STATSTOO) Table 2 for distribution-specific variance.

```{r reduced_models, cache = TRUE}
phylo_only <- brm(Foraged ~ (1 | Species), data = speciesdat, family = bernoulli, cov_ranef = list(phylo = A), cores = 4, control = list(adapt_delta = 0.9))

bayes_R2(phylo_only)

hyp <- paste("sd_Species__Intercept /", "(sd_Species__Intercept + 3.141593^(2/3)) = 0")
(hyp <- hypothesis(phylo_only, hyp, class = NULL))
plot(hyp)

no_phylo <- brm(Foraged ~ DBH + Distance + (1 | Colony), data = speciesdat, family = bernoulli, cores = 4, control = list(adapt_delta = 0.9))
```


## Assessing model fit with and without phylogenetic effects

```{r comparison, cache = TRUE}
loo(complete, phylo_only, no_phylo)
```

We find that the complete model has the best fit to the data

## Diagnostics

We look at the [Pareto k diagnostic values](https://mc-stan.org/loo/reference/pareto-k-diagnostic.html), which should be below 0.7

```{r diagnostics, cache = T}
plot(loo(complete))
```

We see that the estimation is robust.

## Palatability coefficients

UsingWe can look a the coefficients of individual species, which correspond to how likely an they are to be foraged, aftera accounting for all the otehr factors, like the phylogenetic effect.

```{r}

palatcoef <- complete %>% spread_draws(b_Intercept, r_Species[Species ]) %>% group_by(Species) %>% 
  median_qi(Species_mean =  r_Species) %>% mutate(Species = gsub("_"," ", Species))
ggplot(palatcoef, aes(Species_mean)) +geom_histogram() + theme_minimal() 
```

